---
title: "The Data-Type Ontology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Data-Type Ontology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE, comment = "#>")
library(datagoodr)
```

## Why a variable needs more than a type

A CSV cell like `47401` is just text. Is it a ZIP code, a county FIPS, a count
of something? A single "type" label can't capture that: `47401` and `06037` are
both five-digit geographic codes (same *type*), but one is a ZIP and one is a
county (different *meaning*), and neither should be summed like a number
(different *treatment*). The DGF pulls these apart onto separate axes:

| field (DGF column) | question it answers | example values |
|---|---|---|
| **`desired_data_type`** | How should the guide *treat* it? | `number`, `text`, `boolean`, `temporal`, `categorical`, `identifier` |
| **`desired_data_class`** | What *role* does it play? (`class.subclass`) | `mutually_exclusive.geographic`, `id.administrative`, `date.calendar` |
| **`raw_data_format` / `stable_data_format`** | How is it *written*? (a formex expression) | `{{ mmddyyyy }}`, `{{ [strip:$,][dec:.2] }}`, `{{ fips:county }}` |
| **`dd_data_unit`** | At what *unit* is it interpreted? | `USD`, `kg`, `EPSG:4326` |

`type` and `class.subclass` say what a column **is** --- its intended meaning,
stable across file versions. `format` describes how the current file happens to
**write** it, and `dd_data_unit` is a label an analyst needs to interpret it.
That separation is the whole reason the ontology has several fields, not one.

## `data_type` --- the treatment

The top-level type decides how a variable is rendered and summarized. Two of the
groups matter for governance:

**Reader types** --- a plain CSV reader can produce these from the bytes alone:

* **`number`** --- a quantity you can do arithmetic on (revenue, a count, a rate).
* **`text`** --- free or tokenized strings (a name, an email, an address).
* **`boolean`** --- a two-state flag (`Y`/`N`, `TRUE`/`FALSE`, `IN`/`OUT`).
* **`temporal`** --- a point or span in time (a date, a timestamp, a clock time).

**Semantic types** --- meaning a CSV *cannot* carry; only the DGF can assert them:

* **`categorical`** --- a value from a fixed set of groups (a state, an NTEE code,
  a survey response). Stored as text or a number, but *treated* as a category.
* **`identifier`** --- a key that labels a row or entity (an EIN, an ORCID, a
  census tract). Never summarized as a quantity, even when it looks numeric.

(`structured` and `unknown` round out the set as fallbacks.) This
reader-vs-semantic split decides **which bridge** closes the gap between how a
column is stored and how you want it read (see below).

## `data_class` --- the role, with a built-in default

The class is the semantic role, written as **`class.subclass`** (dotted). It is a
two-level hierarchy: the **class** is the broad, looser parent; the **subclass**
refines it and may add its own validation. Name only the class and you get the
parent's treatment; name the subclass to pin the specifics.

| type | classes | example subclasses |
|---|---|---|
| `number` | `quantity`, `currency`, `portion`, `range`, `coordinate` | `count`, `measurement`, `percent`, `lat` |
| `text` | `text`, `name`, `postal`, `web`, `contact`, `color` | `person`, `url`, `email`, `hex` |
| `categorical` | `mutually_exclusive`, `multiselect`, `scale` | `unordered`, `ordered`, `geographic`, `likert` |
| `boolean` | `binary` | `indicator`, `presence`, `status` |
| `temporal` | `date`, `time`, `period`, `phase`, `interval` | `calendar`, `month`, `day_of_week` |
| `identifier` | `id` | `record`, `administrative`, `geographic`, `hash` |

The inheritance has a useful safety property: **defaulting up to the parent can
only relax validation, never falsely reject.** Mark a column
`quantity` but not `count`, and it's validated as a plain number --- you lose the
"non-negative integer" check but never wrongly fail a valid value.

Two columns can share a `type` but differ in `class.subclass` --- a ZIP and an
NTEE code are both `categorical.mutually_exclusive`, but one is `.geographic` and
one is `.classification_code`, and that is what a reader (or an archival standard
like DDI) needs to know.

> This replaces the older `type / subtype / class / format / unit` five-field
> model. The storage hazard the old `numeric_id` / `text_id` / `hash_id`
> subtypes flagged is now handled at the representation layer (`raw_data_type`,
> and ID *hardening* --- see below), not as a semantic coordinate.

## `data_format` --- how it's written (formex)

The format is the surface form the file uses, written as a **formex** expression
inside `{{ }}`. For plain values it's `{{ none }}`; it earns its keep as a
date-mask (`{{ mmddyyyy }}`), a number bundle (`{{ [strip:$,][dec:.2] }}`), a
scheme token (`{{ fips:county }}`), or a directive (`{{ split:[_lo][_hi] }}`).
The **Formex** vignette is the full guide; the short version:

| format | reads |
|---|---|
| `{{ mmddyyyy }}` | a US date, any separator, stabilized to ISO |
| `{{ hh:ii24 }}` | a 24-hour clock time (`hh:ii12A` for 12-hour) |
| `{{ [strip:$,][range:0,Inf][dec:.2] }}` | a dollar amount: strip `$`/`,`, keep 2 decimals |
| `{{ fips:county }}` | a 5-digit county GEOID (registry-backed) |
| `{{ code=label }}` | a coded categorical, `1=Strongly agree` |

## `dd_data_unit` --- the interpretive unit

`dd_data_unit` carries a unit an analyst needs but a type does not: `USD` for a
currency, `kg` for a mass, `EPSG:4326` for a coordinate. It is documentation, not
a parsing instruction.

One thing it deliberately does **not** carry is the temporal *grain*. A date's
grain (calendar vs month vs day-of-week) is load-bearing --- it **selects the
render graphic** --- so it lives in `class.subclass`, where the render engine
reads it:

| `class.subclass` | grain | graphic |
|---|---|---|
| `date.calendar` | a calendar date | calendar heatmap |
| `period.month` | a month bucket | month bars (Jan..Dec) |
| `period.year` | a year | year histogram |
| `phase.hour_of_day` | a time of day | hour-of-day bars |
| `phase.day_of_week` | day of week | Mon..Sun bars |

Retype the grain in the DGF (via `retype_dgf(..., units = c(x = "month"))`, which
writes `period.month`) and re-render --- the chart changes, no re-profiling, because
the frequency table is grain-independent.

## A column reads as type + `class.subclass`

Here is what the migrated `create_dgf()` actually returns for a handful of
columns --- built live from this data:

```{r demo, echo = TRUE, message = FALSE, warning = FALSE}
raw <- data.frame(
  revenue = c(1200.5, 88.0, 45.2, 300.1, 12.0, 7.7),
  state   = c("CA","NY","CA","NY","TX","TX"),
  county  = c("06037","36061","48201","17031","12086","04013"),
  edate   = c("2019-12-31","2020-01-01","2019-06-15","2018-03-02","2021-11-11","2017-07-04"),
  stringsAsFactors = FALSE)
invisible(utils::capture.output(
  dgf <- create_dgf(raw, file = tempfile("dgf"), open = FALSE)))
dgf[, c("var_name", "desired_data_type", "desired_data_class")]
```

`state` and `county` are both geographic, but the **role differs by
cardinality**: `state` repeats, so it's a grouping ---
`categorical.mutually_exclusive.geographic`; the six distinct `county` FIPS read
as a near-unique **key**, so it's promoted to `identifier.id.geographic`. Same
domain, different treatment. (`revenue` carries no detector match, so it keeps
the bare `number` type and its default `quantity` class.)

## The two bridges: `stable` and `desired`

A variable travels from the raw file to how you want to analyze it in **three
stages**, joined by **two rule bridges**:

```
RAW CSV ──raw_to_stable_transform──▶ GOVERNED CSV ──desired_data_import_rule──▶ ANALYSIS
   │            (reformat: lossless)        │           (coerce into type)          │
 raw_*                              stable_*                               desired_*
 how R reads the file               the canonical current state           what it should BE
```

* **`raw_`** --- the file as received (`raw_data_type`, `raw_data_format`).
* **`stable_`** --- the **governed CSV** after cleanup (`stable_data_type`,
  `stable_data_format`) --- the data *as it now is*.
* **`desired_`** --- the intended meaning, unchanging across file versions
  (`desired_data_type`, `desired_data_class`).

Each bridge holds the **name of an `as_*` rule**:

**`raw_to_stable_transform` --- the reformat bridge (→ the stable stage).** A
*lossless* rewrite into a canonical form every tool reads: standardize a weird
date to ISO (`as_mmddyyyy`), strip a currency symbol (`as_usd`), zero-pad and
harden an ID (`as_id[prefix:FIPS]`). Its output is the governed CSV.

**`desired_data_import_rule` --- the import bridge (→ the desired stage).** The
coercion that reads the governed column into the intended type (`as_factor`,
`as_yyyymmdd`, `as_id`). When a value is already unambiguous once stabilized ---
a hardened `FIPS-06037` can only load as text --- the rule is just
`{{ default:character }}`.

### Which bridge? A mechanical rule

The reader-vs-semantic split of `data_type` tells you which bridge closes the gap:

| the desired type is... | example | close the gap with |
|---|---|---|
| **reader** (achievable by a reader) | a date stored as text | **reformat** --- make a plain reader hit it (`raw_to_stable_transform`) |
| **semantic** (not expressible in a CSV) | a ZIP that is `categorical`, an EIN that is an `identifier` | **import rule** --- only metadata can assert it (`desired_data_import_rule`) |

Push as much as possible into the *reformat* --- it is lossless and every
downstream tool benefits --- and reserve import rules for the genuinely
schema-dependent distinctions a CSV can never carry on its own.

## See also

* The **Formex** vignette --- the `{{ }}` format-expression syntax in full, with
  worked examples from the ontology.
* The **Rule-Based Data Governance** vignette --- the three-stage workflow in
  depth, and how `create_dgf()` / `stabilize_data()` / `update_dgf()` use it.
* The **Data-Type Detector Guide** --- the `is_*` predicates that populate these
  coordinates and guard the rules.
* `data-types/research_data_type_ontology_v4.csv` --- the full catalog of
  `type / class / subclass` coordinates, the formex format for each, the
  raw → stable → import workflow by example, and suggested validations.
