---
title: "Rule-Based Data Governance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rule-Based Data Governance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>")
```

## The problem: a CSV has no schema

A CSV is just text. Whether `47401` is a ZIP code, a county FIPS, or a count is
invisible to the reader — every program that opens the file has to *guess*, and
they guess differently. `datagoodr` fixes this by pairing the data with a **Data
Governance File (DGF)**: a small table, one row per variable, that records how
the data is stored, how it should be read, and what it means. The DGF is a
**rule-based** artifact — you write the *name* of a rule in a cell, and the
package applies it — so governance is declarative, portable, and reproducible.

## Three stages, two bridges

A variable's type lives on a short journey from the raw file to how you actually
want to analyze it. The DGF columns are grouped to mirror that journey:

```
RAW CSV ──raw_to_stable_transform──▶ GOVERNED CSV ──desired_data_import_rule──▶ ANALYSIS
   │            (reformat: lossless)        │           (coerce into type)          │
 raw_*                              stable_*                               desired_*
 how R reads the raw file           the canonical current state           what it should BE
 (storage, format)                  (storage, unit, format)               (type, subtype, class)
```

* **`raw_`** — the raw file as received: `raw_data_storage` (how R reads it),
  `raw_data_format`, and profiling (`raw_first5`, `raw_duplicated`).
* **`stable_`** — the **governed CSV** after cleanup: `stable_data_storage`,
  `stable_data_unit`, `stable_data_format`. These describe the data *as it
  currently is*.
* **`desired_`** — the **intended interpretation**, which does not change across
  file versions: `desired_data_type`, `desired_data_subtype`,
  `desired_data_class`.

Between the stages sit two **rule** columns — each holds the *name* of an `as_*`
function:

* **`raw_to_stable_transform`** — the lossless **reformat** (standardize weird
  dates to ISO, strip thousands separators, zero-pad IDs). Its job is to produce
  a *canonical* governed CSV that every tool reads correctly.
* **`desired_data_import_rule`** — the **coercion** that reads the governed
  column into the desired R type (`as_date`, `as_EIN`, ...).

## Which bridge closes the gap? A mechanical rule

The storage a reader produces can be one of only four base types
(`text`, `number`, `boolean`, `temporal`); the *desired* type may also be a
semantic type a CSV cannot express (`categorical`, `identifier`, `list`). That
tells you which bridge to use:

| the desired type is… | example | close the gap with |
|---|---|---|
| **reader-achievable** (base) | a date read as text | **reformat** (`raw_to_stable_transform`) — make the reader hit it |
| **not expressible in a CSV** (semantic) | a ZIP that is a `categorical`, an ID that is an `identifier` | **import rule** (`desired_data_import_rule`) — only metadata can say it |

The design goal is to push as much as possible into the *reformat* (it is
lossless and every downstream tool benefits), leaving import rules only for the
genuinely schema-dependent distinctions.

## `as_*` rules and their `is_*` guards

Every rule is an `as_*` coercion, and every `as_*` has a matching **`is_*`
guard** — and the guards are the detector library documented in the
*Data-Type Detector Guide*. Before `as_ddmmyyyy` runs, `is_ddmmyyyy` checks that
the column actually looks like `dd/mm/yyyy`; if it does not, the rule does not
match the data and you get an error instead of silent corruption. That single
idea ties the whole system together:

* **governance** — drop an `as_*` name in a cell and it is applied;
* **safety** — its `is_*` guard validates the precondition;
* **iteration** — `update_dgf()` re-runs the guards against the *current* data
  and flags rules that no longer fit (because you cleaned the file).

## The workflow

```{r workflow}
library(datagoodr)

# 1. create_dgf() launches the guessers and populates the pipeline.
#    It profiles each column, runs the detector library over it (guess_data_type
#    with the variable name as a hint), and fills in a first pass at every column
#    of the DGF -- storage, a semantic desired_data_type/subtype/class, factor
#    levels, near-unique key flags, and the rg_ profile used to render.
dgf <- create_dgf(raw_data, file = "DGF", open = FALSE)
```

`create_dgf()` does the heavy first guess; **your job is to curate two things**:

**(a) Move the data from raw to canonical** — add reformat rules to
`raw_to_stable_transform` so the governed CSV is clean and self-describing
(standardize dates, declutter grouped numbers, preserve leading zeros). This is
representation only: no information is lost, the values just get written in a
form every tool reads.

**(b) Add semantic context** — set the `desired_*` fields to say what each column
should *be*, independent of how it happens to be stored. A grade coded `1`–`12`
is `categorical/ordinal`, not a number; a FIPS code is `identifier/geographic_id`
when it is a key, `categorical/geography` when it is a grouping variable. This is
where you encode meaning the CSV alone cannot carry.

```{r curate}
# Edit the DGF in Excel, or programmatically:
dgf$raw_to_stable_transform[dgf$var_name == "founded"] <- "as_mmddyyyy"  # -> ISO
dgf$desired_data_type[dgf$var_name == "grade_level"]   <- "categorical"
dgf$desired_data_subtype[dgf$var_name == "grade_level"] <- "ordinal"

# 2. Apply the reformat rules to a raw extract -> a canonical, governed CSV.
stable <- stabilize_data(raw_data, dgf)
write_stable_csv(stable, "data-governed.csv", dgf = dgf)

# 3. Re-profile after you have changed a type or cleaned the data.
dgf <- update_dgf(dgf, raw_data)   # re-runs guards, recomputes rg_* for changed vars
```

## Why the split pays off

* **Iterative improvement.** Governance is not one-shot. You clean a column, and
  `update_dgf()` reconciles the DGF with the new state — the `desired_`
  interpretation is stable, while the `raw_`/`stable_` storage and the rules
  needed to bridge to it evolve as the data improves.
* **Portability.** Because `stable_data_unit`/`stable_data_format` describe the
  data as it is, and the `desired_` fields describe intent, the governed CSV
  carries enough to be understood by non-R tools — and maps cleanly onto
  archival standards (DDI) for deposit in a repository like Dataverse.
* **No hidden magic.** Every transformation is a named rule you can read, and
  every rule has a guard that refuses to run on data it does not fit.

See also the companion *Data-Type Detector Guide* vignette for the `is_*`
detectors that power both `guess_data_type()` and the rule guards.
