---
title: "Formex: the format-expression syntax"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formex: the format-expression syntax}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(datagoodr)
```

Formex is the small language DataGoodr uses to describe, per variable, **how a
raw value is read, cleaned, and imported**. It lives in the machine-read columns
of the DGF (`data_format`, `raw_to_stable_transform`, `stable_format`,
`stable_import_rule`). Every example below is a real row from the type ontology.

## The one rule: `{{ … }}` is the parsed part

A cell is **prose with `{{ }}` interpolations** --- the Mustache/Jinja idiom.
Anything inside `{{ }}` is parsed by Formex; anything outside is a human comment.

```
{{ yyyymmdd }} iso date          # "iso date" is just a comment
{{ [strip:$,][dec:.2] }}
{{ @handle }}                    # leads with { -> Excel can't treat it as a formula
```

Double braces (not single) do two jobs: a regex quantifier is a *single* brace
(`[a-z]{4}`), so `{{ }}` lets single braces appear inside as content; and because
every parsed cell starts with `{`, Excel never mis-reads it as a formula
(`= + - @`), so leading zeros and `@`/`-` survive a CSV round-trip.

A cell holds one of three things:

| kind | looks like | used for |
|---|---|---|
| **formex** | `{{ [strip:$,][dec:.2] }}`, `{{ 9(2)-9(7) }}` | number *bundles* and ID/date *masks* |
| **namespaced token** | `{{ fips:county }}`, `{{ handle:twitter }}` | registry-backed schemes |
| **directive** | `{{ derive }}`, `{{ split:[_lo][_hi] }}`, `{{ none }}` | a mode, not a pattern |

## Reading a row: raw -> stable -> import

Each ontology row tells a small story across five columns. Take
**`number.currency.usd`**:

| column | value | meaning |
|---|---|---|
| `examples` | `$-45 ;; $2300.00 ;; 17,298$` | how the raw value arrives |
| `data_format` | `{{ [strip:$,][range:-Inf,Inf][dec:.2] }}` | the pattern that matches the raw |
| `raw_to_stable_transform` | `{{ as_usd }}` | the rule that cleans it |
| `stable_format` | `{{ as_usd }} -45.00 ;; 2300.00 ;; 17298.00` | the cleaned result |
| `stable_import_rule` | `{{ default:numeric }}` | how to load the stable value |

Strip `$` and `,`, keep two decimals; the result is a plain number, so a normal
read imports it with no extra rule.

A few of these transforms ship today; the fuller `as_*` engine set is
forthcoming. What works now:

```{r transforms}
# same digits: the format's element order decides the date
as_mmddyyyy("03/04/2019")     # US, month-first -> March 4
as_ddmmyyyy("03/04/2019")     # EU, day-first   -> April 3

# identifiers stay text; leading zeros preserved
as_id(c(" 06037 ", "36061"))
```

## Numbers: the four-slot bundle

Number formats are a **bundle of attributes**, `[strip][range][dec][split]`, not
a positional shape. Slots are labeled `[key:value]`, order-free, and empty slots
disappear. `[range:…]` is **checked by validation, never coerced** --- only
`strip` and `dec` change the value.

| row | `data_format` | note |
|---|---|---|
| `count` | `{{ [range:0,Inf][dec:.0] }}` | non-negative integer; import `{{ as_integer }}` (a naive read gives a double) |
| `currency.usd` | `{{ [strip:$,][range:-Inf,Inf][dec:.2] }}` | strip `$` and grouping comma; 2 decimals |
| `currency.euro` | `{{ [strip:€.][range:-Inf,Inf][dec:,2] }}` | European: strip `€` and dot-thousands; **`,2`** = comma is the decimal |
| `portion.percent` | `{{ [strip:%][range:0,100][dec:.n] }}` | strip `%`; bounded 0-100; `.n` = any decimals |

**Ranges split into two variables** (`number.range.min_max`): raw `1.2, 3.4` (or
`20 to 10`), format `{{ lo,hi }}` (or `{{ hi to lo }}`), stable
`{{ split:[_lo][_hi] }}` --- yielding `x_lo` and `x_hi`.

**Coordinates** are numbers with a CRS (`number.coordinate.lat_lon`): raw
`40° 44' 55" N, 74° 00' 22" W`, format `{{ gcs:dms:latlon }}`, transform
`{{ as_coord[from:gcs:dms][order:latlon] }}`, stable
`{{ split:[epsg:4326_lat=40.7128][epsg:4326_lon=-74.006] }}`. `as_coord`
reprojects to WGS-84 decimal degrees and splits.

## Temporal: named-token masks, ISO out

Date/time formats are masks where each token is a semantic unit; doubling sets
width, separators auto-parse, and everything stabilizes to ISO.

| row | raw | `data_format` | stable |
|---|---|---|---|
| `date.calendar` | `03/15/2019 ;; 3-15-2019` | `{{ mmddyyyy }}` | `2019-03-15` |
| `time.clock` | `2:30 PM` | `{{ hh:ii12A }}` | `14:30` |
| `phase.day_of_week` | `Mon` | `{{ WWW }}` | `1` |

Three things to remember:

- **Minute is `ii`, not `mm`** (`mm` is month; `M` is the text month) --- so a
  12-hour clock is `{{ hh:ii12A }}`.
- **Clock convention is an explicit `12`/`24` suffix**, not letter case
  (`hh:ii12` vs `hh:ii24`), so it survives a careless edit. `h`/`H` are accepted
  as synonyms.
- **Phases stabilize to a number** (`Mon -> 1`, week-start declared), stored as
  an ordered factor with the name as the label in `dd_f_levels`.

## Identifiers: masks, schemes, and hardening

An ID format is either a **named scheme** or a **COBOL-picture mask** (`9` digit,
`A` letter, `X` alphanumeric; `(n)` count).

```
record          {{ hash:uuid }}   ->  {{ none }}   ->  import {{ as_id }}
administrative  {{ gov:ein }}     ->  {{ as_ein }} ->  01-2345678 ;; 01-2345678
custom ORG-00123                  ->  {{ A(3)-9(5) }}
```

`as_ein` restores a dropped leading zero *and* the delimiter, so `12345678` and
`01-2345678` land on one canonical value.

**Hardening** (`identifier.id.geographic`): raw `6037 ;; 06037`, format
`{{ fips:county }}`, transform `{{ as_id[prefix:FIPS] }}`, stable
`FIPS-06037`. An all-numeric key is zero-padded **and** prefixed, which forces
every reader to load it as text (note the import is now just
`{{ default:character }}`) and disambiguates a 5-digit FIPS from a 5-digit CBSA.
This is the anti-mojibake default for numeric keys.

## Geographic tokens: `system:unit`

Geographies are a closed, table-backed vocabulary, so the format is a scheme
name: `fips:state` (2), `fips:county` (5), `fips:county3` (the 3-digit
component), `fips:tract` (11), `metro:cbsa`, `postal:zip5` / `postal:zip9`,
`usps:state` (`CA`), `name:state` (`California`). A numeric suffix atomizes a
component; the default is the full GEOID.

## Booleans: presets, unioned

A boolean format is one or more preset encodings joined by `|`, plus a `blank=`
policy (case-insensitive, trimmed). The IRS-990 mixed case:

```
boolean.binary.mixed
  raw     yes ;; 0 ;; t ;; x ;; (blank)
  format  {{ tf|yn|10|check blank=false }}
  stable  TRUE ;; FALSE ;; TRUE ;; TRUE ;; FALSE
```

`boolean.binary.presence` is the checkbox case, `{{ check blank=false }}`. The
`blank=` policy is the one part that can't be defaulted (a blank is *missing* in
a survey, *FALSE* in a checkbox).

## Categoricals: shape + the level dictionary

The format is the encoding *shape*; the code->label mapping lives in
`dd_f_levels`.

| row | `data_format` | note |
|---|---|---|
| `mutually_exclusive.unordered` | `{{ code }}` | `M ;; F` --- labels in `dd_f_levels` |
| `scale.likert` | `{{ code=label }}` | `1=Strongly disagree` --- split code from label |
| `multiselect.tags` | `{{ label [delim:;] }}` | `Volunteer;Donor;Board` |
| `mutually_exclusive.geographic` | `{{ usps:state ;; fips:county }}` | defers to the geo tokens |

Categoricals import with `{{ as_factor }}` / `{{ as_ordered }}`; ordered levels
get their order from the `order` field in `dd_f_levels`.

## Text, and the literal escape

Most text is `{{ none }}` -> `{{ default:character }}`. Some has a scheme:
`web.url` (`{{ url }}` -> `as_url` adds scheme, lowercases host),
`name.handle` (`{{ handle:twitter }}` -> `as_handle` strips `@`, lowercases).

The **literal escape hatch** protects a value from *any* coercion
(`text.text.literal`): format `{{ literal }}`, transform `{{ none }}`, import
`{{ as_text }}`, stable value unchanged. Use it when the raw representation *is*
the data and must not be touched, whatever it looks like. (The `unknown` type
reuses this as its fallback.)

## Directives, escapes, and the format/validation line

| directive | meaning |
|---|---|
| `{{ none }}` | no-op |
| `{{ default:numeric }}` | a plain typed read yields this type; no rule needed |
| `{{ derive }}` | derive the transform from `raw_format -> stable_format` |
| `{{ split:[_lo][_hi] }}` | split one field into two suffixed variables |
| `{{ collapse:[MN=Minnesota] }}` | merge levels (categorical) |
| `{{ re:… }}` | the escape hatch --- full regex for what masks can't express |

`re:` switches to full regex (alternation, lookaround):
`{{ re:\d{3}-\d{2}-\d{4}|\d{2}-\d{7} }}` matches "an SSN or an EIN." A backslash
makes a single character literal; comments live outside the braces.

**Format and validation share the grammar** (a mask shorthand *is* a character
class), so you can tighten a format where it matters:

```
{{ 9(2)-9(7) }}      # parse-sufficient
{{ [0-8](2)-9(7) }}  # same grammar, first two digits constrained
```

Keep the *format* parse-sufficient and let *validation* carry what a mask can't
reach --- a valid-prefix **list**, a checksum, a reference-table lookup. The
nchar-range case (`X(8,10)`) is the interesting middle: both a parse fact and a
constraint, served by one notation.
