---
title: "Guessing a Column's Type"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guessing a Column's Type}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
library(datagoodr)
```

## The chooser above the detectors

The *Data-Type Detector Guide* describes the detector library: dozens of pure
predicates that each answer one yes/no question about a column
(`is_currency_code()`, `is_iso_date()`, `is_county_fips()`, ...).
`guess_data_type()` sits **above** them. Given a column of values, it runs the
whole library, ranks what matched, and returns a single best guess anchored to
the ontology --- the function `create_dgf()` leans on to make its first pass at
typing a dataset.

```{r first}
g <- guess_data_type(c("USD", "EUR", "JPY", "GBP"))
g$guess
g$ontology
```

The guess is a **detector name** (`"currency_code"`); the ontology coordinates
say what that means for governance --- a currency code is a `categorical`
`code`, not a number. The full return value carries the evidence behind the
call:

```{r return}
str(g, max.level = 1)
```

* **`guess`** --- the winning detector, or `NA` when nothing is confident enough.
* **`ontology`** --- `c(data_type, data_subtype, data_class, data_format)` for the
  guess.
* **`confidence`** --- the winner's pass rate (share of values it matched).
* **`candidates`** --- every type that cleared the reporting `floor`, ranked.
* **`n`** --- how many unique values were actually scored.

## How it decides

Four steps, designed to stay cheap on a big dataset:

1. **Dedupe + sample.** Only the unique non-missing values are scored, and at
   most `n_sample` of them (default 100). Scoring a million-row column costs the
   same as scoring its distinct values.
2. **Fast tier first.** The fast detectors run first; the slow tier (fuzzy
   matchers) is consulted only if nothing fast clears `threshold`. (No detectors
   are slow today, so this is currently a single pass.)
3. **Rank with tie-breakers.** Candidates are ordered by pass rate, then by a
   variable-**name** hint, then by **specificity** (a loose, catch-many detector
   loses a tie to a precise one).
4. **Map to the ontology.** The winning detector name is resolved to its
   ontology coordinates.

The raw first step is exposed on its own as `score_data_type()`, which returns
every detector's pass rate against a set of values:

```{r score}
score_data_type(c("USD", "EUR", "XYZ"))[1, ]
```

## The name hint breaks ties the data can't

Some values match more than one type. A five-digit code like `06037` is a valid
ZIP **and** a valid county FIPS --- both detectors pass at 100%. The data alone
cannot choose; the **variable name** can:

```{r namehint}
vals <- c("06037", "36061", "48201", "17031")

guess_data_type(vals, name = "county_fips")$guess
guess_data_type(vals, name = "zip_code")$guess
```

Look at the ranked candidates to see *why*. Both types pass at `1.0`, so the
`name_match` column casts the deciding vote:

```{r candidates}
guess_data_type(vals, name = "county_fips")$candidates
```

With no name supplied, the tie falls to the more **specific** detector, so a
bare column of five-digit codes is not silently mislabeled by whichever type
happens to sort first.

## Confidence, and the `threshold` / `floor` knobs

When nothing matches well, the honest answer is "I don't know":

```{r unknown}
guess_data_type(c("apple", "banana", "cherry"))$guess
```

Two thresholds control how bold the guess is:

* **`threshold`** (default `0.8`) --- the top type must match at least this share
  of values to be *reported* as the guess; below it, `guess` is `NA`.
* **`floor`** (default `0.5`) --- the minimum pass rate for a type to even appear
  in `candidates`.

Lower the threshold to accept a messier column; raise it to demand near-uniform
matches. Because scoring is separated from the decision, you can also inspect
`candidates` yourself and apply your own rule.

```{r knobs}
# a column that is 3/4 ISO dates: unknown at the default, a date if you relax it
messy <- c("2016-03-15", "2016-03-13", "1999-12-31", "not a date")
guess_data_type(messy)$guess               # NA (below 0.8)
guess_data_type(messy, threshold = 0.7)$guess
```

## Where it is used

`create_dgf()` calls `guess_data_type()` once per column (via
`guess_dgf_types()`), passing the variable name as the hint, to seed each
variable's `desired_data_type` / `subtype` / `class` and to flag near-unique
keys. You then correct what the data alone could not reveal --- see the
*Rule-Based Data Governance* vignette for that workflow, and the *Data-Type
Detector Guide* for the predicates underneath.
