---
title: "Research Guide HTML Template"
format:
  html:
    embed-resources: true
editor: visual
---

<br><br><br>

```{r, echo=F}
## pointblank Data Validation Setup

library( pointblank )

agent <- 
  create_agent(
    tbl = small_table,
    tbl_name = "small_table",
    label = "VALID-I Example No. 1"
  ) %>%
  col_is_posix(vars(date_time)) %>%
  col_vals_in_set(vars(f), set = c("low", "mid", "high")) %>%
  col_vals_lt(vars(a), value = 10) %>%
  col_vals_regex(vars(b), regex = "^[0-9]-[a-z]{3}-[0-9]{3}$") %>%
  col_vals_between(vars(d), left = 0, right = 5000) %>%
  interrogate()

html <- 
agent %>%  
  get_agent_report() %>%
  htmltools::as.tags()
```

## Demo Data Dictionary

<br><br>

::: parent
::: div1
#### VarName_01_X
:::

::: div2
Ipsum is simply dummy text of the printing and typesetting industry.
:::

::: div3
|            |             |
|------------|-------------|
| DATA TYPE: | **numeric** |
| SCOPE:     | **PZ**      |
| LENGTH:    | **6**       |
:::

::: div4
DEFINITION: Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.
:::
:::

<br>

::: parent
::: div1
#### VarName_02_L2
:::

::: div2
Ipsum is simply dummy text of the printing and typesetting industry.
:::

::: div3
|            |            |
|------------|------------|
| DATA TYPE: | **factor** |
| SCOPE:     | **PC**     |
| LENGTH:    | **24**     |
:::

::: div4
DEFINITION: Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.

##### LEVELS:

|       |        |
|-------|:-------|
| **A** | goat   |
| **B** | monkey |
| **C** | turtle |
| **D** | horse  |
:::
:::

```{r, eval=F, echo=F}
::: div5
##### LEVELS:

|       |        |
|-------|:-------|
| **A** | goat   |
| **B** | monkey |
| **C** | turtle |
| **D** | horse  |
:::
:::
```

<br>

::: parent
::: div1
#### VarName_03_with_length_up_to_32
:::

::: div2
Ipsum is simply dummy text of the printing and typesetting industry.
:::

::: div3
|            |             |
|------------|-------------|
| DATA TYPE: | **numeric** |
| SCOPE:     | **EZ**      |
| LENGTH:    | **13**      |
:::

::: div4
DEFINITION: Lorem Ipsum is simply dummy text of the printing and typesetting industry. Lorem Ipsum has been the industry's standard dummy text ever since the 1500s, when an unknown printer took a galley of type and scrambled it to make a type specimen book.
:::

::: div5
```{r, results='asis', echo=F}
html  # add pointblank validation table
```
:::
:::

<link href='https://fonts.googleapis.com/css?family=Fira Code' rel='stylesheet'> <link rel="stylesheet" type="text/css" href="//fonts.googleapis.com/css?family=Anonymous+Pro" />

```{=html}
<style>

@import url('https://fonts.cdnfonts.com/css/aharoni');

.parent {
    display: grid;
    grid-template-columns: repeat(3, 1fr);
    grid-template-rows: auto;  /* repeat(2, 0.2fr) 1fr 1fr; */
    grid-column-gap: 20px;
    grid-row-gap: 10px;
    grid-template-areas:
      "a a a"
      "b b b"
      "c d d"
      "e e e" ;
}

/* grid-row-start / grid-column-start / grid-row-end / grid-column-end */
/* .div1 { grid-area: 1 / 1 / 2 / 4; } */
/* .div2 { grid-area: 2 / 1 / 3 / 4; } */
/* .div3 { grid-area: 3 / 1 / 4 / 2; } */
/* .div4 { grid-area: 3 / 2 / 4 / 4; } */


.div1 { grid-area: a; }

.div2 { 
    grid-area: b;
    padding-left: 50px;
}

.div3 { 
    grid-area: c;
    padding-left: 50px;
}

.div4 { 
    grid-area: d;
    padding-left: 20px;
}

.div5 { 
    grid-area: e;
    padding-left: 50px;
}

h1.title { color: lightgray; }
h2.anchored { color: lightgray; }
h3.anchored { color: gray; }

h4 {
    text-decoration: underline;
    text-underline-offset: 10px;
    text-decoration-thickness: 2px;
    font-family: Georgia, Times, "Times New Roman", serif; 
    /* font-family: 'Fira Code'; */
    /* font-family: 'Aharoni', sans-serif; */
    font-size: 2.2em;
    font-weight: 700;  
    /* line-height: 120px; */ 
}

h5 { font-size: 0.9em; }

.div2 p { 
    font-family: Verdana;
    font-size: 1.1em;
}

.div3 td { 
    /* font-family: "Anonymous Pro"; */
    /* font-family: Georgia, Times, "Times New Roman", serif; */ 
    /* font-weight: 400; */
    font-family: 'Aharoni', sans-serif;
    font-size: 0.9em;
}

.div4 p { 
    font-family:"Calibri", sans-serif;
    font-size: 0.8em;
}

.div4 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div4 table {
    margin-left: 20px;
}

.div5 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div5 table {
    margin-left: 20px;
}

.table>tbody {
    border-top: none;
}

.table>:not(caption)>*>* {
    padding: 0rem 0rem;
}

tbody, tfoot, tr, td, th {
    border-color: inherit;
    border-style: none; 
    border-width: 0;
}


#pb_agent .gt_table {
    border-top-width: 0px; 
    /* border-top-color: #A8A8A8; */
}  

#pb_agent .gt_title {
    font-size: 1.3em !important;
}

#pb_agent .gt_left { color: black !important; }

</style>
```
<br><br>

<hr>

<br><br>

# HTML

<br>

### Code

    <div class="div1"><h4>VarName_01_X</h4></div>
    <div class="div2">
      Ipsum is simply dummy text of the printing and typesetting industry. 
    </div>
    <div class="div3">
      <table>
        <tr>
    <td>DATA TYPE</td>
    <td>numeric</td>
        </tr>
        <tr>
    <td>SCOPE</td>
    <td>PZ</td>
        </tr>
        <tr>
    <td>LENGTH</td>
    <td>6</td>
        </tr>
        </table>
    </div>
    <div class="div4">
      DEFINITION: Lorem Ipsum is simply dummy text of the printing and typesetting 
      industry. Lorem Ipsum has been the industry's standard dummy text ever since 
      the 1500s, when an unknown printer took a galley of type and scrambled it to 
      make a type specimen book.
    </div>

<br>

<hr>

<br>

### CSS

    <style>

    @import url('https://fonts.cdnfonts.com/css/aharoni');

    .parent {
        display: grid;
        grid-template-columns: repeat(3, 1fr);
        grid-template-rows: auto;  
        grid-column-gap: 20px;
        grid-row-gap: 10px;
        grid-template-areas:
          "a a a"
          "b b b"
          "c d d"
          "e e e" ;
    }

    /* grid-row-start / grid-column-start / grid-row-end / grid-column-end */
    /* .div1 { grid-area: 1 / 1 / 2 / 4; } */
    /* .div2 { grid-area: 2 / 1 / 3 / 4; } */
    /* .div3 { grid-area: 3 / 1 / 4 / 2; } */
    /* .div4 { grid-area: 3 / 2 / 4 / 4; } */


    .div1 { grid-area: a; }

    .div2 { 
        grid-area: b;
        padding-left: 50px;
    }

    .div3 { 
        grid-area: c;
        padding-left: 50px;
    }

    .div4 { 
        grid-area: d;
        padding-left: 20px;
    }

    .div5 { 
        grid-area: e;
        padding-left: 50px;
    }

    h1.title { color: lightgray; }
    h2.anchored { color: lightgray; }
    h3.anchored { color: gray; }

    h4 {
        text-decoration: underline;
        text-underline-offset: 10px;
        text-decoration-thickness: 2px;
        font-family: Georgia, Times, "Times New Roman", serif; 
        font-size: 2.5em;
    }

    h5 { font-size: 0.9em; }

    .div2 p { 
        font-family: Verdana;
        font-size: 1.1em;
    }

    .div3 td { 
        font-family: 'Aharoni', sans-serif;
        font-size: 0.9em;
    }

    .div4 p { 
        font-family:"Calibri", sans-serif;
        font-size: 0.8em;
    }

    .div4 td { 
        font-family: "Anonymous Pro";
        font-size: 0.9em;
    }

    .div4 table {
        margin-left: 20px;
    }

    .div5 td { 
        font-family: "Anonymous Pro";
        font-size: 0.9em;
    }

    .div5 table {
        margin-left: 20px;
    }

    .table>tbody {
        border-top: none;
    }

    .table>:not(caption)>*>* {
        padding: 0rem 0rem;
    }

    tbody, tfoot, tr, td, th {
        border-color: inherit;
        border-style: none; 
        border-width: 0;
    }

    </style>
