---
title: "Research Guide"
format:
  html:
    embed-resources: true
editor: visual
---

<br><br><br>

```{r, warning=F, message=F, include=F}
library( htmltools )
library( dplyr )
library( pointblank )
library( knitr )
library( gt )
library( kableExtra )
library( psych )
```

```{r, echo=F}

# source functions 
# import DGF 
# read data
```

DGF (dd) - vname - vtype - vlabel - vdesc - ftable - scope - loc

```{r, results="asis"}

for( i in 1:nrow(dd) )
{
  vname <- dd$vname[i]
  v <- d[[vname]]
  vtype <- dd$vtype[i]
  vlabel <- dd$vlabel[i]
  vdesc <- dd$vdesc[i]
  scope <- dd$scope[i]
  loc <- dd$loc[i]
  
  ftable <- NULL
  
  if( vtype == "factor" )
  { ftable <- get_ftable( dd$ftable[i] ) }
  
  print_v( 
    vname=vname,
    v=v,
    vtype=vtype,
    vlabel=vlabel,
    vdesc=vdesc,
    scope=scope,
    loc=loc,
    ftable=ftable )
}

cat( "{{< pagebreak >}} \n\n")

```

<br>

<hr>

<br>

```{r, eval=F}
#| echo: false

get_section <- function( row.name )
{
  ds <- dplyr::filter( dfg, var_name== row.name )
  create_section( x1=ds$desc, x2=ds$k, x3=ds$loc )
}

print_dd <- function( dfg, vars=c("r2","r2","r3"), group.name=NULL )
{
   # for( i in vars )
   # {
   #   get_section( row.name=i )
   # }
   dfg2 <- dplyr::filter( dfg, var_name %in% vars )
   sapply( dfg2, get_section )
}
```

<link href='https://fonts.googleapis.com/css?family=Fira Code' rel='stylesheet'> <link rel="stylesheet" type="text/css" href="//fonts.googleapis.com/css?family=Anonymous+Pro" />

````{=html}
<style>

@import url('https://fonts.cdnfonts.com/css/aharoni');

.parent {
    display: grid;
    grid-template-columns: repeat(3, 1fr);
    grid-template-rows: auto;  /* repeat(2, 0.2fr) 1fr 1fr; */
    grid-column-gap: 20px;
    grid-row-gap: 10px;
    grid-template-areas:
      "a a a"
      "b b b"
      "c d d"
      "e e e" ;
}

/* grid-row-start / grid-column-start / grid-row-end / grid-column-end */
/* .div1 { grid-area: 1 / 1 / 2 / 4; } */
/* .div2 { grid-area: 2 / 1 / 3 / 4; } */
/* .div3 { grid-area: 3 / 1 / 4 / 2; } */
/* .div4 { grid-area: 3 / 2 / 4 / 4; } */


.div1 { grid-area: a; }

.div2 { 
    grid-area: b;
    padding-left: 50px;
}

.div3 { 
    grid-area: c;
    padding-left: 50px;
}

.div4 { 
    grid-area: d;
    padding-left: 20px;
}

.div5 { 
    grid-area: e;
    padding-left: 50px;
}

h1.title { color: lightgray; }
h2.anchored { color: lightgray; }
h3.anchored { color: gray; }

h4 {
    text-decoration: underline;
    text-underline-offset: 10px;
    text-decoration-thickness: 2px;
    font-family: Georgia, Times, "Times New Roman", serif; 
    /* font-family: 'Fira Code'; */
    /* font-family: 'Aharoni', sans-serif; */
    font-size: 2.2em;
    font-weight: 700;  
    /* line-height: 120px; */ 
}

h5 { font-size: 0.9em; }

.div2 p { 
    font-family: Verdana;
    font-size: 1.1em;
}

.div3 td { 
    /* font-family: "Anonymous Pro"; */
    /* font-family: Georgia, Times, "Times New Roman", serif; */ 
    /* font-weight: 400; */
    font-family: 'Aharoni', sans-serif;
    font-size: 0.9em;
}

.div4 p { 
    font-family:"Calibri", sans-serif;
    font-size: 0.8em;
}

.div4 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div4 table {
    margin-left: 20px;
}

.div5 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div5 table {
    margin-left: 20px;
}

.table>tbody {
    border-top: none;
}

.table>:not(caption)>*>* {
    padding: 0rem 0rem;
}

tbody, tfoot, tr, td, th {
    border-color: inherit;
    border-style: none; 
    border-width: 0;
}


#pb_agent .gt_table {
    border-top-width: 0px; 
    /* border-top-color: #A8A8A8; */
}  

#pb_agent .gt_title {
    font-size: 1.3em !important;
}

#pb_agent .gt_left { color: black !important; }

</style>





```{=html}
<style>

@import url('https://fonts.cdnfonts.com/css/aharoni');

.parent {
    display: grid;
    grid-template-columns: repeat(3, 1fr);
    grid-template-rows: auto;  /* repeat(2, 0.2fr) 1fr 1fr 1fr; */
    grid-column-gap: 20px;
    grid-row-gap: 10px;
    grid-template-areas:
      "a a a"
      "b b b"
      "c d d"
      "e f g" 
      "h i i";
}



.div1 { grid-area: a; }

.div2 { 
    grid-area: b;
    padding-left: 50px;
}

.div3 { 
    grid-area: c;
    padding-left: 50px;
}

.div4 { 
    grid-area: d;
    padding-left: 20px;
}

.div5 { 
    grid-area: e;
    padding-left: 50px;
}

.div6 { 
    grid-area: f;
    padding-left: 50px;
}

.div7 { 
    grid-area: g;
    padding-left: 50px;
}

.div8 { 
    grid-area: h;
    padding-left: 50px;
    display: grid;
}

.div8 .cell-output-display { 
    display: grid;
}

.div8 p { 
    display: grid;
}


.div9 { 
    grid-area: i;
    padding-left: 20px;
}

h1.title { color: lightgray; }
h2.anchored { color: lightgray; }
h3.anchored { color: gray; }

h4 {
    text-decoration: underline;
    text-underline-offset: 10px;
    text-decoration-thickness: 2px;
    font-family: Georgia, Times, "Times New Roman", serif; 
    /* font-family: 'Fira Code'; */
    /* font-family: 'Aharoni', sans-serif; */
    font-size: 1.8em;
    font-weight: 700;  
    /* line-height: 120px; */ 
}

h5 { font-size: 0.9em; }

.div2 p { 
    font-family: Verdana;
    font-size: 1.1em;
}

.div3 td { 
    /* font-family: "Anonymous Pro"; */
    /* font-family: Georgia, Times, "Times New Roman", serif; */ 
    /* font-weight: 400; */
    font-family: 'Aharoni', sans-serif;
    font-size: 0.9em;
}

.div4 p { 
    font-family:"Calibri", sans-serif;
    font-size: 0.8em;
}

.div4 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div4 table {
    margin-left: 20px;
}

.div5 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div5 table {
    margin-left: 20px;
}

.div6 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div6 table {
    margin-left: 20px;
}

.div7 td { 
    font-family: "Anonymous Pro";
    /* font-weight: 400; */
    font-size: 0.9em;
}

.div7 table {
    margin-left: 20px;
}

.table {
    width: 80%;
}

.table>tbody {
    border-top: none;
}

.table>:not(caption)>*>* {
    padding: 0rem 0rem;
}

tbody, tfoot, tr, td, th {
    border-color: inherit;
    border-style: none; 
    border-width: 0;
}


.div5 tr {
  border-bottom: 1px solid;
  border-color: #D3D3D3;
}

.div5 th {
  display:none;
}

.div6 tr {
  border-bottom: 1px solid;
  border-color: #D3D3D3;
}

.div6 th {
  display:none;
}

.div7 tr {
  border-bottom: 1px solid;
  border-color: #D3D3D3;
}

.div7 th {
  display:none;
}

.div8 img, svg {
    /* vertical-align: bottom;*/
    align-self: end;
    justify-self: end;
}

#pb_agent .gt_table {
    border-top-width: 0px; 
    /* border-top-color: #A8A8A8; */
}  

#pb_agent .gt_title {
    font-size: 1.3em !important;
}

#pb_agent .gt_left { color: black !important; }


@media print {
   body {
   display:table;
   table-layout:fixed;
   padding-top:2.5cm;
   padding-bottom:3cm;
   padding-left:1.5cm;
   padding-right:1.5cm;
   height:auto;
    }
}


</style>
````
